---
title: "Data Analysis"
output:
  html_document: default
  pdf_document: default
---

_Note: Write R code in external R scripts, and use `read_chunk()` to read them into the current document._

### Initial Evironment Set-up

The tidyverse package set.

```{r, echo = FALSE}
knitr::read_chunk('../data_analysis/code/install_packages.r')
```

```{r, packages, echo = FALSE}
```

### Acquire the Surgical Wait Times Data

Obtain the surgical wait times data from the Nova Scotia Government [Open Data Portal](https://data.novascotia.ca).

- Navigate to the Open Data Portal
- Select the Export option
- Select [CSV](https://data.novascotia.ca/api/views/wu5w-qxki/rows.csv?accessType=DOWNLOAD)
- Save the data file to the project subdirectory (data_analysis/data).

_The data set has been manually updated by Carlo Carandang per the comments in the data analytics report. A detailed algorithm addressing the various strategies to impute the missing values for some features remains to be documented for reproducibility as newer data becomes available._

### Load the Surgical Wait Times Data

```{r, echo = FALSE}
knitr::read_chunk('../data_analysis/code/load_data.r')
```

```{r, dataset, echo = FALSE}
```

### Wrangle the Data

```{r, echo = FALSE}
knitr::read_chunk('../data_analysis/code/wrangle_data.r')
```

```{r, wrangling, echo = FALSE}
```

Determine the feature names.

```{r, column-names, echo = FALSE, comment = ''}
```

Examine the data set for any missing values.

```{r, missing-values-check, echo = FALSE, comment = ''}
```

```{r, echo = FALSE}
knitr::read_chunk('../data_analysis/code/visualize_data.r')
```

```{r, feature-relationships, echo = FALSE}
```

### Build Model

```{r, echo = FALSE}
knitr::read_chunk('../data_analysis/code/build_model.r')
```

A linear regression model is constructed.

```{r, linear-regression-model, echo = FALSE, results = 'hide'}
```

```{r, prediction-model, echo = FALSE, comment = ''}
``` 

```{r, anova-coefficients, echo = FALSE, results = 'markup'}
```
