---
title: "Improving a Multiple Regression Model by Choosing a Different Baseline Coefficient"
author: "Carlo Carandang"
date: "November 23, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Machine Learning and Multiple Linear Regression

We recently used machine learning, specifically multiple linear regression, in a surgical wait times data set, where the surgical specialties were the independent (feature) variables, and the predicted surgical wait time the dependent (outcome) variable. We decided to blog about this analysis, as the choice of the default B0 coefficient, cardiac surgery, had different implications for significance of the other individual coefficients than choosing a different default B0 coefficient, such as general surgery. 

## Cardiac Surgery as the Default B0 Coefficient

Cardiac surgery was chosen as the default B0 coefficient by R application as it ordered the independent variables in alphabetical order, with cardiac surgery being the first in the alphabetized list. The following analysis shows the summary statistics of the features, where cardiac surgery is picked by R as the default B0 coefficient:

```{r, echo = FALSE}
## ---- linear-regression-model ----
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(stringr)
library(tibble)
library(broom)
library(purrr)
library(e1071)
library(magrittr)

wait_times <- read.delim2(file = '../data_analysis/data/Surgical_Wait_Times_2017-09-04.csv')

feature_name <- wait_times %>% names()
feature_name %>% noquote()

names(wait_times) <- wait_times %>% names() %>% tolower()

wait_times <- wait_times %>%
  mutate(specialty = tolower(specialty),
         specialty = str_replace(specialty, " surgery", ""),
         procedure = tolower(procedure))

specialty_type <- wait_times %>% filter(!is.na(specialty)) %>% select(specialty) %>% 
  distinct() %>% arrange(specialty) %>%  as.data.frame()
#specialty_type %>% print.data.frame(right = FALSE, row.names = FALSE)

data_frame(
  feature = c('specialty', 'procedure'),
  missing_count = c(wait_times %>% filter(is.na(specialty)) %>% tally(),
                    wait_times %>% filter(is.na(procedure)) %>% tally()) %>% unlist(),
  nonmissing_count = c(wait_times %>% filter(!is.na(specialty)) %>% tally(),
                       wait_times %>% filter(!is.na(procedure)) %>% tally()) %>% unlist()
) %>% arrange(feature) #%>% print.data.frame(right = FALSE, row.names = FALSE)

wait_times %>% group_by(procedure) %>% filter(!is.na(procedure)) %>% tally() %>%
  arrange(desc(n)) %>% rename('observations' = 'n') %>% 
  as.data.frame() %>% head(5) #%>% 
  #print.data.frame(right = FALSE, row.names = FALSE)

#isolate rows with procedure = 'all'
wait_times <- wait_times %>% filter(procedure == "all")

as.data.frame(
  cbind(
    t(wait_times %>% 
        select(everything()) %>% 
        summarise_all(funs(sum(is.na(.))))), 
    t(wait_times %>% select(everything()) %>% 
        summarise_all(funs(sum(!is.na(.))))))) %>% 
  rownames_to_column() %>% 
  mutate(feature = rowname, missing_count = V1, nonmissing_count = V2) %>% 
  transmute(feature, missing_count, nonmissing_count) %>% arrange(feature) #%>%
  #print.data.frame(right = FALSE, row.names = FALSE)

complete_observation_count <- sum(wait_times %>% complete.cases())

wait_time_by_specialty <- wait_times %>% 
  filter(!is.na(consult_90th) & !is.na(surgery_90th)) %>%
  select(specialty, consult_90th, surgery_90th) %>% 
  group_by(specialty) %>% 
  summarise(minimum = as.integer(min(consult_90th + surgery_90th, na.rm = TRUE)), 
            maximum = max(consult_90th + surgery_90th, na.rm = TRUE), 
            average = as.integer(median(consult_90th + surgery_90th, na.rm = TRUE)),
            sigma = as.integer(sd(consult_90th + surgery_90th, na.rm = TRUE)),
            total = as.integer(sum(consult_90th + surgery_90th, na.rm = TRUE)),
            observations = n())

as.data.frame(wait_time_by_specialty) #%>% 
  #print.data.frame(right = FALSE, row.names = FALSE)


#isolate rows with procedure = 'all'
wait_times <- wait_times %>% filter(procedure == "all")

#remove rows with null values in Consult_90th column
wait_times <- wait_times[!is.na(wait_times$consult_90th),]

#remove rows with null values in Surgery_90th column
wait_times <- wait_times[!is.na(wait_times$surgery_90th),]

#A bivariate linear regression model is constructed with two dependent variables (consult_90th and surgery_90th), representing the 90th percentiles for each instance of a surgical specialty’s wait time, added together to give the combined surgical wait time and one independent variable.
specialty_factor <- wait_times %>% select(specialty) %>% flatten_chr()

wait_times <- wait_times %>% 
  mutate(specialty =  specialty_factor)

specialty_consult90 <- wait_times %>% select(consult_90th) %>% unlist()
specialty_surgery90 <- wait_times %>% select(surgery_90th) %>% unlist()
specialty90 <- specialty_consult90 + specialty_surgery90
specialty <- wait_times %>% select(specialty) %>% unlist()

model <- lm(formula = specialty90 ~ specialty)


## ---- model ----
model
summary(model)
glance(model)

statistics <- glance(model)

k <- statistics %>% select(df) %>% first()
n <- statistics %>% select(df.residual) %>% first()
p <- statistics %>% select(p.value)
rse <- statistics %>% select(sigma)

f <- statistics %>% select(statistic)
f_critical <- qf(0.95, k, n - k - 1)
f_critical
```
## General Surgery as the Default B0 Coefficient

General surgery was chosen as the default B0 coefficient to observe if it had any changes in the in statistical significance of the other independent variables, when comapred to choosing cardiac surgery as the default. The following analysis shows the summary statistics of the features, where general surgery is picked as the default B0 coefficient:

```{r, echo = FALSE}
## ---- linear-regression-model ----
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(stringr)
library(tibble)
library(broom)
library(purrr)
library(e1071)
library(magrittr)

wait_times <- read.delim2(file = '../data_analysis/data/Surgical_Wait_Times_2017-09-04.csv')

feature_name <- wait_times %>% names()
feature_name %>% noquote()

names(wait_times) <- wait_times %>% names() %>% tolower()

wait_times <- wait_times %>%
  mutate(specialty = tolower(specialty),
         specialty = str_replace(specialty, " surgery", ""),
         procedure = tolower(procedure))

specialty_type <- wait_times %>% filter(!is.na(specialty)) %>% select(specialty) %>% 
  distinct() %>% arrange(specialty) %>%  as.data.frame()
#specialty_type %>% print.data.frame(right = FALSE, row.names = FALSE)

data_frame(
  feature = c('specialty', 'procedure'),
  missing_count = c(wait_times %>% filter(is.na(specialty)) %>% tally(),
                    wait_times %>% filter(is.na(procedure)) %>% tally()) %>% unlist(),
  nonmissing_count = c(wait_times %>% filter(!is.na(specialty)) %>% tally(),
                       wait_times %>% filter(!is.na(procedure)) %>% tally()) %>% unlist()
) %>% arrange(feature) #%>% print.data.frame(right = FALSE, row.names = FALSE)

wait_times %>% group_by(procedure) %>% filter(!is.na(procedure)) %>% tally() %>%
  arrange(desc(n)) %>% rename('observations' = 'n') %>% 
  as.data.frame() %>% head(5) #%>% 
  #print.data.frame(right = FALSE, row.names = FALSE)

#isolate rows with procedure = 'all'
wait_times <- wait_times %>% filter(procedure == "all")

as.data.frame(
  cbind(
    t(wait_times %>% 
        select(everything()) %>% 
        summarise_all(funs(sum(is.na(.))))), 
    t(wait_times %>% select(everything()) %>% 
        summarise_all(funs(sum(!is.na(.))))))) %>% 
  rownames_to_column() %>% 
  mutate(feature = rowname, missing_count = V1, nonmissing_count = V2) %>% 
  transmute(feature, missing_count, nonmissing_count) %>% arrange(feature) #%>%
  #print.data.frame(right = FALSE, row.names = FALSE)

complete_observation_count <- sum(wait_times %>% complete.cases())

wait_time_by_specialty <- wait_times %>% 
  filter(!is.na(consult_90th) & !is.na(surgery_90th)) %>%
  select(specialty, consult_90th, surgery_90th) %>% 
  group_by(specialty) %>% 
  summarise(minimum = as.integer(min(consult_90th + surgery_90th, na.rm = TRUE)), 
            maximum = max(consult_90th + surgery_90th, na.rm = TRUE), 
            average = as.integer(median(consult_90th + surgery_90th, na.rm = TRUE)),
            sigma = as.integer(sd(consult_90th + surgery_90th, na.rm = TRUE)),
            total = as.integer(sum(consult_90th + surgery_90th, na.rm = TRUE)),
            observations = n())

as.data.frame(wait_time_by_specialty) #%>% 
  #print.data.frame(right = FALSE, row.names = FALSE)

#isolate rows with procedure = 'all'
wait_times <- wait_times %>% filter(procedure == "all")

#remove rows with null values in Consult_90th column
wait_times <- wait_times[!is.na(wait_times$consult_90th),]

#remove rows with null values in Surgery_90th column
wait_times <- wait_times[!is.na(wait_times$surgery_90th),]

#Prior to building the statistical model the baseline factor is ‘general surgery’ instead of the default ‘cardiac surgery’ to determine the impact, if any, on the linear regression model w.r.t. the null hypothesis.

#A bivariate linear regression model is constructed with two dependent variables (consult_90th and surgery_90th), representing the 90th percentiles for each instance of a surgical specialty’s wait time, added together to give the combined surgical wait time and one independent variable.
specialty_factor <- wait_times %>% select(specialty) %>% flatten_chr() %>% 
  as.factor() %>% relevel('general')

wait_times <- wait_times %>% 
  mutate(specialty =  specialty_factor)

specialty_consult90 <- wait_times %>% select(consult_90th) %>% unlist()
specialty_surgery90 <- wait_times %>% select(surgery_90th) %>% unlist()
specialty90 <- specialty_consult90 + specialty_surgery90
specialty <- wait_times %>% select(specialty) %>% unlist()

model <- lm(formula = specialty90 ~ specialty)


## ---- model ----
model
summary(model)
glance(model)

statistics <- glance(model)

k <- statistics %>% select(df) %>% first()
n <- statistics %>% select(df.residual) %>% first()
p <- statistics %>% select(p.value)
rse <- statistics %>% select(sigma)

f <- statistics %>% select(statistic)
f_critical <- qf(0.95, k, n - k - 1)
f_critical
```

## Effect of choosing a different baseline B0 coefficient

When comparing the summary statistics of choosing cardiac surgery versus general surgery as the default B0 coefficient, we do see a difference in the number of individual coefficients being statistically significant at a confidence level of 95%: cardiac surgery as default has 4 statistically significant coefficients, while general surgery as default has 5 statistically significant coefficients. 

However, when looking at the statistical significance of the overall model, the F-statistic for both baselines are equal, at F-statistic = 1.76. Therefore, the choice of the default B0 coefficient has no effect on the statistical significance of the overall model.

## Multiple Inference and Interpreting Multiple Coefficents

When interpreting more than one coefficient in a regression equation, it is important to use appropriate methods for multiple inference, rather than using just the individual confidence intervals that are automatically given by most software. One technique for multiple inference in regression is using confidence regions.
https://www.ma.utexas.edu/users/mks/statmistakes/regressioncoeffs.html
https://www.ma.utexas.edu/users/mks/statmistakes/multipleinference.html
